---
title: "In Class Exercises - Sorting and Filtering"
author: "Jenna Ahart"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

## Setup and Loading Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the packages we'll need
# Remember, think of this like packing for our trip
# (We don't take the whole closet, only what we need)
library(tidyverse)
library(lubridate)

```

Run this line below load the data for this assignment. It will load a dataframe called "impeach," containing U.S. House Democrats and whether they supported the first impeachment of Pres. Trump, along with when they announced such support.

```{r}

impeach <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRh8d5JaDqBtByzNw2JZF3idaACobDhhk-p7chJoktA0HawELHFOvjQOqCpzGA4MGONvPlR7GASqW-K/pub?gid=1765341510&single=true&output=csv")


```

Let's see what we have:

```{r}

impeach


```

## Exercise Questions

*For the questions below, write your R code to return the requested records in the code chunk provided.*

1)  The column "for_impeachment" indicates whether the member has publicly called for an impeachment inquiry. Filter to return only the ones where the answer is NO.

**I used the filter command here to select only House members who answered "NO" for for_impeachment**

```{r}

filter(impeach, for_impeachment == "NO")





```

2)  Filter to return only results where a member is both against impeachment, and comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column).

**I combined commands by taking my last command, and inserting a comma and the new condition for p16winningparty as R to signify Trump won the district in 2016.**

```{r}



filter(impeach, for_impeachment == "NO", p16winningparty == "R")


```

3)  Filter for only results where a member is against impeachment, comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column), and also comes from a district that Mitt Romney won in 2012 ("p12winningparty").

**Again, I took my command from the previous question and added to it so that p12winningparty must also be R for Romney.**

```{r}

filter(impeach, for_impeachment == "NO", p16winningparty == "R", p12winningparty == "R")




```

4)  Filter for only results from September 2019 where a member is a YES for impeachment.

**I used the same format for my commands as before, this time using for_impeachment as yes instead of no. I also added the conditions that date_month is 9 and date_year is 2019. I could have inserted a range of possible date_comb values instead, but I found this method more straight forward.**

```{r}

filter(impeach, for_impeachment == "YES", date_month == "9", date_year == "2019")



```

5)  Filter for only results where a member is a YES for impeachment and is from a district where Clinton won more than 70 percent of the vote in 2016 (found in column "clinton_percent").

**I used the same format for my command, but I changed used the "\>" sign so that all the clinton_percent values must be greater than 70.**

```{r}


filter(impeach, for_impeachment == "YES", clinton_percent > 70)



```

6)  Sort the entire dataframe based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

**I used the arrange command this time, sorting by pct_bachelors. This automatically put the data in ascending order based on the outlined criteria.**

```{r}



arrange(impeach, pct_bachelors)



```

7)  Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

**I used %\>% this time to combine commands. I made for_impeachment NO and sorted for pct_bachelors using parts of previous lines.**

```{r}

impeach %>%
  filter(for_impeachment == "NO")%>%
  arrange(pct_bachelors)
  



```

8)  Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

Then filter those records by only those whose bachelor's percentage is below the national average (found in the "pct_bachelors_compared_to_national" column).

**I used the same command as before, adding the line that pct_bachelors_compared_to_national must be "BELOW" to meet the final criteria.**

```{r}

impeach %>%
  filter(for_impeachment == "NO")%>%
  arrange(pct_bachelors)%>%
  filter(pct_bachelors_compared_to_national == "BELOW")




```

9)  Filter for only members from New Jersey who are NO on impeachment.

**I created a command to filter the data so that the state is NJ and the impeachment answer is NO. There was only one match.**

```{r}

filter(impeach, state == "NJ", for_impeachment == "NO")




```

10) Filter for those who were YES on impeachment, with a declared date prior to 2019. So only those with dates before 2019. Then sort those so that the highest Clinton vote percentages are at the top.

**I used %\>% to combine commands again. I started by making for_impeachment must be YES. Then, I filtered so that the date_year must be before (or less than 2019). Finally, I arranged based on clinton_percent. I used the "desc" command to make sure the values were in descending order.**

```{r}

impeach %>%
  filter(for_impeachment == "YES")%>%
  filter(date_year < 2019) %>%
  arrange(desc(clinton_percent))
  



```
